---
title: "Profit & Loss"
subtitle: "Year-to-Date"
output: word_document
params:
  CANexchange: NA
  EUROexchange: NA
  income: NA
  expenses: NA
  dates: NA
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(knitr)
```

**Report Period:** `r params$dates[1]` to `r params$dates[2]`

*Report Generated:* `r Sys.Date()`

This report includes income and expense tracking from all business accounts n all currencies (US, CAN, and EURO).  Income is reported based on payment date to align with tax liability.  Expenses are reported based on withdrawal date (checking) or transaction date (credit).

This document serves an example of how to render a custom report from Shiny.

# Gross Income

```{r grossIncomeTotals, include=FALSE}
incData <- as.data.frame(params$income) 
totUsaInc <-  sum(incData$PayAmount[incData$PayCurrency=="US"], na.rm=TRUE)
totCanInc <-  sum(incData$PayAmount[incData$PayCurrency=="CAN"], na.rm=TRUE)
totEuroInc <-  sum(incData$PayAmount[incData$PayCurrency=="EURO"], na.rm=TRUE)

# Total gross calculated in US dollars based on exchange rate entered in Shiny.
grossInc <- totUsaInc + (totCanInc * params$CANexchange) + (totEuroInc * params$EUROexchange)

```

**`r paste("$", formatC(grossInc, big.mark=",", format = "d"), "US dollars")`**

 - *Total US Income:* `r paste("$",formatC(totUsaInc, big.mark=",", format = "d"))`

 - *Total CAN Income:* `r paste("$",formatC(totCanInc, big.mark=",", format = "d"))`

 - *Total EURO Income:* `r paste("$",formatC(totEuroInc, big.mark=",", format = "d"))`


```{r incomeTables, echo=FALSE}

# Profit in cash accounting is only money received, so remove pending payments
bizInc <- incData %>%
	dplyr::filter(IncType %in% c("Consulting", "Teaching")) %>%
	dplyr::arrange(PayCurrency, PayDate) %>%
	dplyr::select(PayDate, IncType, Client, InvNum, PayAmount, PayCurrency) %>%
	dplyr::filter(!is.na(PayDate))

otherInc <- incData %>%
	dplyr::filter(!IncType %in% c("Consulting", "Teaching")) %>%
	dplyr::arrange(PayCurrency, PayDate) %>%
	dplyr::select(PayDate, IncType, PayAmount, PayCurrency) %>%
	dplyr::filter(!is.na(PayDate))
```

## Summary of Income from Contracted Work

```{r contractSummary, echo=FALSE}
contractSummary <- bizInc %>%
	dplyr::group_by(PayCurrency, Client) %>%
	dplyr::summarize(Total = sum(PayAmount), .groups = 'drop')

knitr::kable(contractSummary)
```

## Summary of Income from Other Sources

```{r otherIncomeSummary, echo=FALSE}
otherSummary <- otherInc %>%
	dplyr::group_by(PayCurrency, IncType) %>%
	dplyr::summarize(Total = sum(PayAmount), .groups = 'drop')
knitr::kable(otherSummary)
```


# Expenses

```{r expenseData, include=FALSE}
expData <- as.data.frame(params$expenses)

reportClasses <- params$expenses %>%
	dplyr::select(ExpClass, ExpCategory) %>%
	dplyr::distinct() %>%
	dplyr::arrange(ExpClass, ExpCategory)
	
# Total by currency
spendUs <- expData %>%
	dplyr::filter(ExpCurrency == "US") %>%
	dplyr::group_by(ExpClass, ExpCategory) %>%
	dplyr::summarize(TotalUs = round(sum(ExpAmount),2), .groups = 'drop')

spendCan <- expData %>%
	dplyr::filter(ExpCurrency == "CAN") %>%
	dplyr::group_by(ExpClass, ExpCategory) %>%
	dplyr::summarize(TotalCan = round(sum(ExpAmount),2), .groups = 'drop')

spendTotal <- left_join(reportClasses, spendUs, by = c("ExpClass", "ExpCategory")) %>%
	left_join(spendCan, by = c("ExpClass", "ExpCategory")) %>%
	dplyr::mutate(TotalUs = ifelse(is.na(TotalUs), 0, TotalUs),
								TotalCan = ifelse(is.na(TotalCan), 0, TotalCan),
								exchCanToUs = round(TotalCan * params$CANexchange),
								`TOTAL (US)` = TotalUs + exchCanToUs) %>%
	dplyr::select(-exchCanToUs)

# Extract out member draw, because for a sole member LLC this is not an expense - but keep to report
# Extract out transfers and perdiem expenses, if present, because transfers are not an expense and per diem is calculated elsewhere
memberDrawTotal <- dplyr::filter(spendTotal, ExpClass == "Member Draw")
spendTotal <- dplyr::filter(spendTotal, !ExpClass %in% c("Member Draw", "Transfer"), !ExpCategory %in% c("Per Diem Expense"))

```

```{r expenseTotals, include=FALSE}
totUsaExp <-  sum(spendTotal$TotalUs)
totCanExp <-  sum(spendTotal$TotalCan)

totalExp <- sum(spendTotal$`TOTAL (US)`)
```

**`r paste("$", formatC(totalExp, big.mark=",", format = "d"), "US dollars")`**

 - *Total US Expenses:* `r paste("$",formatC(totUsaExp, big.mark=",", format = "d"))`

 - *Total CAN Expenses:* `r paste("$",formatC(totCanExp, big.mark=",", format = "d"))`


## Summary of Expenses (Grouped by Class)

```{r expenseSummary, echo=FALSE}
expenseSummary <- spendTotal %>%
	dplyr::group_by(ExpClass) %>%
	dplyr::summarize(`TOTAL (US)` = sum(`TOTAL (US)`), .groups = 'drop')

knitr::kable(expenseSummary)
```

# Net Income

```{r netINcome, echo=FALSE}
netInc <- grossInc - totalExp
totalMember <- memberDrawTotal$`TOTAL (US)`
bizProfits <- netInc - totalMember

```


**`r paste("$",formatC(netInc, big.mark=",", format = "d"), "US dollars")`**

Of the net income, `r paste("$",formatC(totalMember, big.mark=",", format = "d"), "US dollars")` was transferred to personal accounts as Member Draws.

The final business balance for this period, after subtracting Member Draws, is `r paste("$",formatC(bizProfits, big.mark=",", format = "d"), "US dollars")`.

*Note that all final values in US dollars are near estimates as the exchange rate is applied to income and expenses that remain in the original currency.  Estimates here are solely for the purposes of estimating business value and tax liability in US currency.*

*Note also that this code is for demonstration purposes only and any financial paperwork should be reviewed by an accountand and tax professional.*

# Appendix

## Business Contract Income

```{r earnedIncome, echo=FALSE}
knitr::kable(bizInc)
```

## Other Income

```{r otherIncome, echo=FALSE}
knitr::kable(otherInc)
```

## Expenses (Grouped by Category)

```{r allExpenses, echo=FALSE}
knitr::kable(spendTotal)
```
